---
title: "Air Quality and Wildfire Trends Over the Last 5 Years"
author: "Alani Gayles (aig6@uw.edu), Shruti Anand (shrana3@uw.edu), An Dao (andao27@uw.edu), Brandon Mochizuki (bmoch@uw.edu)"
date: "11/14/2022"
output: html_document
---

```{r setup, include=FALSE}
source("../source/data_access.R")
source("../source/chart_acres_burned.R")
source("../source/chart_aqi_states.R")
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis

### Chart 1: Choropleth

```{r echo=FALSE}
choropleth
```
In order to see which states have been hit hardest by wildfires, we decided to
make a choropleth that maps each state with the total acres burned by wildfires
from 2017 to 2021. This chart is color encoded by acres burned. White represents
0 acres burned while dark red represents 8 million acres, which is fitting since
red is associated with fire. States can be hovered for detailed information
such as state name and the totals for every year. As shown, California has
the most acres burned by far, and generally the west coast seems to be more
prone to wildfire burnage, including Alaska.

### Chart 2: Scatter Plot
```{r echo=FALSE}
scatter
```
In order to see if there is a relationship between wildfires and AQI, we decided
to make a scatter plot that graphs total burned acres with average AQI in 2021 for
all states. A scatterplot was used since trends between two variables can be
spotted. If there is a correlation, there should be some sort of line that shows
up. Since not all states had wildfires in 2021, only the ones that had are
graphed. Hovering over a point shows the state and the specific numbers. As shown,
while there are strong cases like California with the worst AQI and most acres
burned, most of the remaining states are fairly inconclusive.

### Chart 3: Line
