---
title: "Air Quality and Wildfire Trends Over the Last 5 Years"
author: "Alani Gayles (aig6@uw.edu), Shruti Anand (shrana3@uw.edu), An Dao (andao27@uw.edu), Brandon Mochizuki (bmoch@uw.edu)"
date: "11/14/2022"
output: html_document
---

```{r setup, include=FALSE}
source("../source/data_access.R")
source("../source/sum_info.R")
source("../source/table.R")
source("../source/chart_acres_burned.R")
source("../source/chart_aqi_states.R")
source("../source/chart_wa_aqi_2021.R")
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis

### Dataset Summary

The datasets used in this project are a dataset of fires from MTBS and
datasets of AQI from 2017 to 2021 from CBSA. The fires dataset contains
information about `r sum_info$num_wildfires` wildfires from 2017 to 2021. The
AQI dataset has `r sum_info$num_observations_aqi` AQI observations of
states in the U.S. from 2017 to 2021. The average national AQIs for every year
are `r sum_info$average_AQI`, ordered from 2017 to 2021. TODO Complete summary
paragraph with two other variables from sum_info.R.

### Table

```{r echo = FALSE, results = "asis"}
knitr::kable(table, caption = "Total acres burned every year in states")
```

This table shows the amount of acres burned in every state, sorted by total acres
burned. We included this table since we are trying to see if there is a correlation
between wildfires and AQI, so it's useful to be aware of wildfire hot spots.

TODO Add "why you included the table, and what information it reveals
(California has most, states next to Pacific Ocean in general have more, etc).

### Chart 1: Choropleth

```{r echo=FALSE}
choropleth
```

In order to see which states have had the most acres burned by wildfires, we
decided to make a choropleth that maps each state with the total acres burned by
wildfires from 2017 to 2021. This chart is color encoded by acres burned. White
represents 0 acres burned while dark red represents 8 million acres, which is
fitting since red is associated with fire. States can be hovered for detailed
information such as state name and the totals for every year. As shown,
California has the most acres burned by far, and generally the west coast seems
to be more prone to wildfire burnage, including Alaska.

### Chart 2: Scatter Plot

```{r echo=FALSE}
scatter
```

In order to see if there is a relationship between wildfires and AQI, we decided
to make a scatter plot that graphs total burned acres with average AQI in 2021 for
all states. A scatterplot was used since trends between two variables can be
spotted. If there is a correlation, there should be some sort of line that shows
up. Since not all states had wildfires in 2021, only the ones that had are
graphed. Hovering over a point shows the state and the specific numbers. As shown,
while there are strong cases like California with the worst AQI and most acres
burned, most of the remaining states are fairly inconclusive.

### Chart 3: Line

```{r echo=FALSE}
chart
```

In order to see the AQI in Washington, we decided to make a line chart of the AQI
in 2021. TODO Fill out "why you included the chart and what information it reveals,
including notable observations and insights from the chart."