---
title: "Air Quality and Wildfire Trends Over the Last 5 Years"
author: "Alani Gayles (aig6@uw.edu), Shruti Anand (shrana3@uw.edu), An Dao (andao27@uw.edu), Brandon Mochizuki (bmoch@uw.edu)"
date: "11/14/2022"
output: html_document
---

```{r setup, include=FALSE}
source("../source/data_access.R")
source("../source/sum_info.R")
source("../source/table.R")
source("../source/chart_acres_burned.R")
source("../source/chart_aqi_states.R")
source("../source/chart_wa_aqi_2021.R")
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis

### Dataset Summary

The datasets used in this project are a dataset of fires from MTBS and
datasets of AQI from 2017 to 2021 from CBSA. The fires dataset contains
information about `r sum_info$num_wildfires` wildfires from 2017 to 2021. The
AQI dataset has `r sum_info$num_observations_aqi` AQI observations of
states in the U.S. from 2017 to 2021. The average national AQIs for every year
are `r sum_info$average_AQI`, ordered from 2017 to 2021. TODO Complete summary
paragraph with two other variables from sum_info.R.

### Table

```{r echo = FALSE, results = "asis"}
knitr::kable(table, caption = "Total acres burned every year in states")
```

For our table, we wanted to see how much acres were burned in every state.
Each sorted by the total amount of acres burned. To do this, we had to get all the 
wildfire data and the corresponding state within the timeline of 2017 to 2021. There 
were a few "states" that had to be dropped because they weren't actually in the United 
States. This table is included because we wanted to see if there was a correlation 
between the wildfires and AQI in each state. This is because it will be easier to spot 
any wildfire hot spots in the United States. It is shown that California had the most 
wildfires and highest AQI. In correlation, states along the Pacific Ocean tend to also 
have a higher amount of wildfires and levels of AQI. 

### Chart 1: Choropleth

```{r echo=FALSE}
choropleth
```

In order to see which states have had the most acres burned by wildfires, we
decided to make a choropleth that maps each state with the total acres burned by
wildfires from 2017 to 2021. This chart is color encoded by acres burned. White
represents 0 acres burned while dark red represents 8 million acres, which is
fitting since red is associated with fire. States can be hovered for detailed
information such as state name and the totals for every year. As shown,
California has the most acres burned by far, and generally the west coast seems
to be more prone to wildfire burnage, including Alaska.

### Chart 2: Scatter Plot

```{r echo=FALSE}
scatter
```

In order to see if there is a relationship between wildfires and AQI, we decided
to make a scatter plot that graphs total burned acres with average AQI in 2021 for
all states. A scatterplot was used since trends between two variables can be
spotted. If there is a correlation, there should be some sort of line that shows
up. Since not all states had wildfires in 2021, only the ones that had are
graphed. Hovering over a point shows the state and the specific numbers. As shown,
while there are strong cases like California with the worst AQI and most acres
burned, most of the remaining states are fairly inconclusive.

### Chart 3: Line

```{r echo=FALSE}
chart
```

In order to see the AQI in Washington, we decided to make a line chart of the AQI each 
day in 2021. The use of this chart helps us see the pattern of AQI throughout the
year in Washingtonand we can see that there was a large spike around the time of 
August and September when wildfireseason is the worst. Over the summer, there is drier 
heat, less rain, and more factors that contibuteto high levels of air quality. However, 
we can see that air quality remains consistent throughtout the rest of the year with 
no significantly noticable spikes. 

